Push-Pull Signal-Function Functional Reactive Programming

### 2012-10-03 Docs/IFL2012/presentation/presentation.tex

\begin{frame}
\frametitle{Functional Reactive Programming (FRP)}
    \begin{itemize}
        \item {\em Functional} First-class and higher-order functions.
        \item {\em Reactive} Behavior changes in response to temporal inputs.
        \item Basic abstractions:
        \begin{itemize}
            \item Signals - Functions of time\footnote{Often also called behaviors.}.
            \item Events - Sequences of temporally ordered and labeled discrete values.
        \end{itemize}
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Signal-Function FRP}
    \begin{itemize}
        \item Signal functions are transformers of events and signals.
        \item Signal functions are first class in Signal-Function FRP.
        \item Signals and events are not first-class in Signal Function FRP.
        \item This approach is more composable than first-class signals and events (since input may be transformed)\footcite{Courtney2001-1}.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Push vs. Pull Evaluation}
    \begin{itemize}
        \item When to evaluate what?
        \item Pull evaluation (``demand-driven''):\\ Evaluate when output is needed.
        \item Push evaluation (``data-driven''):\\ Evaluate when input is available.
        \item Ideally, an FRP system uses push for events and pull for signals\footcite{Elliott2009}.
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Separating Events and Signals}
    \begin{itemize}
        \item Necessary to lift any function to a corresponding signal function, without input/output type annotation
        \item Events encoded as optional signals:
\begin{Verbatim}
type Event a = Maybe a
\end{Verbatim}
        \item Combined signals and events encoded as one signal with pair type:
    \end{itemize}
\end{frame}

@@

    \begin{itemize}
        \item Described by Sculthorpe\footcite{Sculthorpe2011}
        \item Describe separation of individual signals and events
        \item In Haskell (with {\tt -XEmptyDataDecls}):
\begin{Verbatim}
data SVEmpty
data SVSignal a
data SVEvent a
data SVAppend svl svr
type :^: svl svr = SVAppend svl svr
\end{Verbatim}
        \item Would like to use {\tt -XDataKinds} but it's not working well yet.
    \end{itemize}

@@

% Add conclusion

\begin{frame}
\frametitle{Further Work}
\begin{itemize}
  \item Dynamic collections: Dynamically switch between collections of signal functions.
  \item Semantics/correctness proof (especially for event merging).
  \item Time-independence optimization.
\end{itemize}
\end{frame}

>>>>

\begin{frame}
\frametitle{Functional Reactive Programming (FRP)}
    \begin{itemize}
        \item {\em Functional}: First-class and higher-order functions.
        \item {\em Reactive}: Behavior changes in response to temporal inputs.
        \item Basic abstractions:
        \begin{itemize}
            \item Signals - Functions of time\footnote{Often also called behaviors.}.
            \item Events - Sequences of temporally ordered and labeled\\
                  discrete values.
        \end{itemize}
    \item ``Classic'' FRP - Signals and events are first-class.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Signal-Function FRP (SF-FRP)}
    \begin{itemize}
        \item Signal functions are transformers of events and signals.
        \item Signal functions are first-class in Signal-Function FRP.
        \item Signals and events are not first-class in Signal-Function FRP.
        \item This approach is more composable than first-class\\
              signals and events (since input may be transformed)\footcite{Courtney2001-1}.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Push vs. Pull Evaluation}
    \begin{itemize}
        \item When to evaluate what?
        \item Pull evaluation (``demand-driven''):\\ Evaluate when output is needed.
        \item Push evaluation (``data-driven''):\\ Evaluate when input is available.
        \item Ideally, an FRP system uses push for events and\\
              pull for signals\footcite{Elliott2009}.
        \item This has been achieved for ``classic'' FRP
              (first class signals and events) but not SF-FRP\footcite{Elliott2009}.
    \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Separating Events and Signals}
    \begin{itemize}
        \item The abstraction must lift any function to a corresponding signal function, without input/output type annotation.
        \item Events encoded as optional signals:
\begin{Verbatim}
type Event a = Maybe a
\end{Verbatim}
        \item Combined signals and events encoded as one signal with\\
              a pair type:
    \end{itemize}
\end{frame}

@@

    \begin{itemize}
        \item Described by Sculthorpe\footcite{Sculthorpe2011}.
        \item Describe separation of individual signals and events.
        \item In Haskell (with {\tt -XEmptyDataDecls}):
\begin{Verbatim}
data SVEmpty
data SVSignal a
data SVEvent a
data SVAppend svl svr
type :^: svl svr = SVAppend svl svr
\end{Verbatim}
        \item Would like to use {\tt -XDataKinds} but it's not working well yet.
    \end{itemize}

@@

\begin{frame}
\frametitle{Conclusion}
    \begin{itemize}
        \item Signal vectors fix an abstraction leak in Signal-Function FRP.
        \item Representing events distinctly from signals in the type permits
              heterogeneous (push-pull) evaluation of signal functions.
    \end{itemize}
\end{frame}

\begin{frame}
\frametitle{Further Work}
    \begin{itemize}
        \item Dynamic collections: Dynamically switch between collections of signal functions.
        \item Demonstrations and performance comparisons.
        \item Semantics/correctness proof (especially for event merging).
        \item Time-independence optimization.
    \end{itemize}
\end{frame}
